% Paper A research brief for July 2026, Week 3.
% Date range: Mon 2026-07-20 to Sun 2026-07-26.
% Status snapshot: Thu 2026-07-23 10:59 PT.

\begin{frame}
  \titlepage
  \vspace{-1.0em}
  \begin{center}
    \small Paper A research brief: July 2026, Week 3 \\
    \scriptsize Status through Thu 2026-07-23 10:59 PT
  \end{center}
\end{frame}

\begin{frame}{Research question and method}
  \begin{block}{Question}
    When can an agent answer from a short learned state, and when should it return to raw evidence?
  \end{block}
  \begin{enumerate}\small
    \item Recurrently encode context chunks into soft memory while carrying earlier memory forward.
    \item Read memory through a small LoRA adapter on the same frozen language model.
    \item Keep an adapter-free bounded raw path on the same backbone.
    \item Use held-out confidence to study routing between the two paths.
  \end{enumerate}
  \begin{block}{Positioning}
    \footnotesize Compression ratio alone is not the contribution. Paper A studies the complete
    \textbf{compact-memory versus raw-evidence interface}, including paired failure, coverage, and cost.
  \end{block}
\end{frame}

\begin{frame}{Core result: useful compact state, clear boundary}
  \centering\scriptsize
  \begin{tabular}{llrrr}
  \toprule
  base & method & QuALITY & BFCL & HotpotQA \\
  \midrule
  Qwen3-8B & \textcolor{gray}{Raw ref.} & --$^\dagger$ & 92.4 & 53.7 \\
            & \textcolor{gray}{SFT ref.} & --$^\dagger$ & 95.4 & 68.8 \\
            & LLMLingua-2 & --$^\dagger$ & 70.3 & 22.1 \\
            & Compressor (w/o gate) & --$^\dagger$ & 72.3 & 28.9 \\
            & \textbf{Compressor (w/ gate)} & --$^\dagger$ & \textbf{88.5} & \textbf{50.9} \\
  \midrule
  Qwen3.5-9B & \textcolor{gray}{Raw ref.} & --$^\dagger$ & 84.5 & 53.9 \\
              & \textcolor{gray}{SFT ref.} & --$^\dagger$ & 94.9 & 71.7 \\
              & LLMLingua-2 & --$^\dagger$ & 60.8 & 28.9 \\
              & Compressor (w/o gate) & --$^\dagger$ & 72.0 & 30.5 \\
              & \textbf{Compressor (w/ gate)} & --$^\dagger$ & \textbf{80.5} & \textbf{51.7} \\
  \bottomrule
  \end{tabular}
  \vspace{0.45em}
  \begin{itemize}\footnotesize
    \item $^\dagger$QuALITY is quarantined; BFCL and HotpotQA remain valid.
    \item SFT is a full-cost adaptation reference, not a compression competitor.
    \item Formal testing certifies no nonzero coverage; routing is reported as held-out empirical analysis.
  \end{itemize}
\end{frame}

\begin{frame}{Real long-context transfer}
  \centering\scriptsize
  \begin{tabular}{llrrr}
  \toprule
  base & method & LongBench-v2 & $\infty$Bench-choice & BABILong \\
  \midrule
  Qwen3-8B & \textcolor{gray}{Raw ref.} & 31.2 & 52.8 & 18.1 \\
            & Compressor (w/o gate) & --$^\dagger$ & --$^\dagger$ & --$^\ddagger$ \\
            & \textbf{Compressor (w/ gate)} & --$^\dagger$ & --$^\dagger$ & \textbf{18.1} \\
  Qwen3.5-9B & \textcolor{gray}{Raw ref.} & 33.0 & 52.4 & 10.5 \\
              & Compressor (w/o gate) & --$^\dagger$ & --$^\dagger$ & --$^\ddagger$ \\
              & \textbf{Compressor (w/ gate)} & --$^\dagger$ & --$^\dagger$ & \textbf{11.9} \\
  \bottomrule
  \end{tabular}
  \vspace{0.45em}
  \begin{block}{Task-structured operating region}
    \footnotesize The compressor without the gate can beat truncated raw on multi-hop LongBench transfer, while raw wins on
    single-document / extractive targets. The gate selects between the two paths:
    \textbf{Compressor (w/ gate) $\ge$ Raw on every harvested two-base target}.
  \end{block}
  \footnotesize $^\dagger$QuALITY-source invalid; $^\ddagger$BABILong collapse. RULER is in an OOM loop.
\end{frame}

\begin{frame}{Execution status and next steps}
  \begin{columns}[T]
  \begin{column}{0.56\textwidth}
    \centering\scriptsize
    \begin{tabular}{lrrrr}
    \toprule
    stage & done & run & fail & pending \\
    \midrule
    core main & 60 valid & 0 & 28 invalid & 0 \\
    replicate + SFT audit & 0 valid & 0 & 9 invalid & 0 \\
    transfer adapters & 27 valid & 0 & 16 invalid & 0 \\
    real long-context & 81 valid & 0 & 25 affected & 0 \\
    generality & 13 valid & 2 invalid & 8 & 6 \\
    budget / RULER & 0 valid & 1 window & 2 fail & 18 \\
    ablation & 0 & 0 & 0 & 36 \\
    official baselines & 0 & 0 & 0 & 52 \\
    \bottomrule
    \end{tabular}
  \end{column}
  \begin{column}{0.44\textwidth}
    \textbf{Current workers}
    \begin{itemize}\scriptsize
      \item 2 stale-loader xLAM QuALITY runs;
      \item RULER window-8192 is running;
      \item primary GPU 1 is contended.
    \end{itemize}
    \textbf{Order to completion}
    \begin{enumerate}\scriptsize
      \item stop stale/OOM workers;
      \item sync fixes and validate;
      \item rerun quarantined cells;
      \item repair RULER/K32;
      \item ablation, cost, baselines.
    \end{enumerate}
  \end{column}
  \end{columns}
\end{frame}
