% July Week 3 weekly: Mon 2026-07-20 to Sun 2026-07-26.
% Status snapshot: Thu 2026-07-23 10:59 PT.

\providecommand{\paref}[2]{\href{../../papers/paper-a-latent-memory/#1}{\textcolor{blue}{#2}}}

\section{July Week 3 (Mon 2026-07-20 to Sun 2026-07-26)}

\begin{frame}{Reporting Period}
  \begin{center}
    \textbf{Plan 08 / Long-Context Compression --- July Week 3} \\[0.3em]
    \small Mon 2026-07-20 to Sun 2026-07-26 \\
    \scriptsize Status through Thu 2026-07-23 10:59 PT. \\
    \scriptsize Current priority: finish \textbf{Paper A} with complete long-context, generality,
    budget, ablation, cost, and official-baseline evidence.
  \end{center}
\end{frame}

\begin{frame}{Paper A --- Introduction logic (six paragraphs)}
  \begin{columns}[T]
  \begin{column}{0.49\textwidth}
    \begin{enumerate}\small
      \item \textbf{Why long context matters.}\\
      It is the working state of document, tool, conversation, and agent systems.
      \vspace{0.5em}
      \item \textbf{The problem is not only capacity.}\\
      Long inputs cost more, but even a context that fits may be used unreliably.
      \vspace{0.5em}
      \item \textbf{Compression is not automatically robust.}\\
      A shorter state may remove the fact or relation required by the next decision.
    \end{enumerate}
  \end{column}
  \begin{column}{0.51\textwidth}
    \begin{enumerate}\setcounter{enumi}{3}\small
      \item \textbf{The overlooked question is failure detection.}\\
      Average compressed accuracy does not say whether memory is sufficient for this input.
      \vspace{0.5em}
      \item \textbf{Agent design should preserve two paths.}\\
      Use compact memory when trusted; let a gate return to raw evidence otherwise.
      \vspace{0.5em}
      \item \textbf{GCM implements the interface.}\\
      One frozen backbone builds recurrent memory, reads it, estimates confidence, and provides raw fallback.
    \end{enumerate}
  \end{column}
  \end{columns}
  \begin{block}{Paper focus}
    \footnotesize \textbf{Robust long-context compression}: map where soft memory works, detect where it does
    not, and measure how much a shared-backbone fallback recovers.
  \end{block}
\end{frame}

\begin{frame}{Paper A --- story and completed core evidence}
  \small
  \begin{block}{Story}
    Gated Context Memory (GCM) treats soft memory as a \textbf{routed interface}: use a recurrent
    short state when it is sufficient, and preserve a bounded raw path when the needed evidence does not
    survive compression.
  \end{block}
  \centering\scriptsize
  \begin{tabular}{llrrr}
  \toprule
  base & method & QuALITY & BFCL & HotpotQA \\
  \midrule
  Qwen3-8B & Raw ref. & --$^\dagger$ & 92.4 & 53.7 \\
            & LLMLingua-2 & --$^\dagger$ & 70.3 & 22.1 \\
            & Compressor (w/o gate) & --$^\dagger$ & 72.3 & 28.9 \\
            & \textbf{Compressor (w/ gate)} & --$^\dagger$ & \textbf{88.5} & \textbf{50.9} \\
  Qwen3.5-9B & Raw ref. & --$^\dagger$ & 84.5 & 53.9 \\
              & LLMLingua-2 & --$^\dagger$ & 60.8 & 28.9 \\
              & Compressor (w/o gate) & --$^\dagger$ & 72.0 & 30.5 \\
              & \textbf{Compressor (w/ gate)} & --$^\dagger$ & \textbf{80.5} & \textbf{51.7} \\
  \bottomrule
  \end{tabular}
  \vspace{0.35em}
  \begin{itemize}\footnotesize
    \item $^\dagger$QuALITY is quarantined after a zero-based label-index bug; BFCL/Hotpot remain valid.
    \item Formal fixed-family test certifies no nonzero coverage; routing remains held-out empirical analysis.
  \end{itemize}
\end{frame}

\begin{frame}{Paper A --- real long-context result}
  \centering\scriptsize
  \begin{tabular}{llrrr}
  \toprule
  base & method & LongBench-v2 & $\infty$Bench-choice & BABILong \\
  \midrule
  Qwen3-8B & Raw ref. & 31.2 & 52.8 & 18.1 \\
            & Compressor (w/o gate) & --$^\dagger$ & --$^\dagger$ & --$^\ddagger$ \\
            & \textbf{Compressor (w/ gate)} & --$^\dagger$ & --$^\dagger$ & \textbf{18.1} \\
  Qwen3.5-9B & Raw ref. & 33.0 & 52.4 & 10.5 \\
              & Compressor (w/o gate) & --$^\dagger$ & --$^\dagger$ & --$^\ddagger$ \\
              & \textbf{Compressor (w/ gate)} & --$^\dagger$ & --$^\dagger$ & \textbf{11.9} \\
  \bottomrule
  \end{tabular}
  \vspace{0.4em}
  \begin{block}{Finding}
    \footnotesize On multi-hop transfer, the compressor without the gate can beat truncated raw
    (Qwen3 HotpotQA $19.1\to30.5$; Qwen3.5 $3.9\to26.6$).
    On single-document / extractive targets, the gated variant falls back.
    Across the harvested two-base targets, \textbf{Compressor (w/ gate) $\ge$ Raw in every cell}.
  \end{block}
  \footnotesize $^\dagger$QuALITY-source invalid; $^\ddagger$Hotpot$\to$BABILong collapse.
  \paref{paper-a-all-benchmarks-skeleton-2026-07-21.md}{all-benchmark tables}
\end{frame}

\begin{frame}{Paper A --- full status and execution order}
  \begin{columns}[T]
  \begin{column}{0.56\textwidth}
    \centering\scriptsize
    \begin{tabular}{lrrrr}
    \toprule
    stage & done & run & fail & pending \\
    \midrule
    core main & 60 valid & 0 & 28 invalid & 0 \\
    replicate + SFT audit & 0 valid & 0 & 9 invalid & 0 \\
    transfer adapters & 27 valid & 0 & 16 invalid & 0 \\
    real long-context & 81 valid & 0 & 25 affected & 0 \\
    7-model generality & 13 valid & 2 invalid & 8 & 6 \\
    budget / RULER & 0 valid & 1 window & 2 fail & 18 \\
    ablation & 0 & 0 & 0 & 36 \\
    official baselines & 0 & 0 & 0 & 52 \\
    \bottomrule
    \end{tabular}
  \end{column}
  \begin{column}{0.44\textwidth}
    \textbf{Live}
    \begin{itemize}\scriptsize
      \item 2 stale-loader xLAM QuALITY workers;
      \item RULER raw window-8192 is running;
      \item primary GPU 1 is memory-contended.
    \end{itemize}
    \textbf{Next}
    \begin{enumerate}\scriptsize
      \item stop stale/OOM workers;
      \item sync corrected loader and code;
      \item validate one clean QuALITY cell;
      \item rerun quarantined cells;
      \item repair RULER, then ablation/baselines.
    \end{enumerate}
  \end{column}
  \end{columns}
  \vspace{0.4em}
  \centering\footnotesize
  \paref{paper-a-full-benchmark-table-2026-07-21.md}{benchmark + baseline inventory}
  \quad
  \paref{experiment-receipt-2026-07-16.md}{experiment receipt}
\end{frame}
