\documentclass[aspectratio=169]{beamer}
\usetheme{Madrid}
\usecolortheme{default}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\hypersetup{hidelinks}
\usepackage{xcolor}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, calc}

% ---- GitHub hyperlink helpers (repo: BDeMo/research-notes) ----
\newcommand{\ghbase}{https://github.com/BDeMo/research-notes/blob/main/notes/plans/08-compressed-context-memory}
\newcommand{\ghB}[2]{\href{\ghbase/papers/paper-b-forgetting-gating/#1}{\color{blue}\underline{#2}}}
\newcommand{\ghC}[2]{\href{\ghbase/papers/paper-c-linear-attention/#1}{\color{blue}\underline{#2}}}
\newcommand{\ghA}[2]{\href{\ghbase/papers/paper-a-latent-memory/#1}{\color{blue}\underline{#2}}}
\newcommand{\best}[1]{\textbf{#1}}
\newcommand{\refm}[1]{\textcolor{black!55}{#1}}

\title[Long-Context Memory]{Long-Context Compression \& Linear-Attention Memory}
\subtitle{Research progress --- Papers A / B / C}
\author{Mingjia Shi}
\institute{RCA Foundation Model / Long-Context Memory}
\date{\today}

\begin{document}

\begin{frame}\titlepage
  \begin{center}\scriptsize All artifacts hyperlinked to
  \href{https://github.com/BDeMo/research-notes}{\color{blue}\underline{github.com/BDeMo/research-notes}} (blue = clickable).\end{center}
\end{frame}

% ============================ INDEX ============================
\begin{frame}{Outline / Index}
  \begin{columns}[T]
  \begin{column}{0.5\textwidth}
  \textbf{Part I --- Framing}
  \begin{enumerate}\small
    \item Background: the long-context memory wall
    \item One thesis, three papers (the mainline)
  \end{enumerate}
  \vspace{0.6em}
  \textbf{Part II --- Paper B (parallel line)}
  \begin{enumerate}\setcounter{enumi}{2}\small
    \item The diagnosis (model-invariant)
    \item The method: IMP router
    \item Full results table (all methods $\times$ benches)
    \item Honest standing vs \emph{all} baselines
    \item Rigor: a bug we found \& fixed
    \item Occam simplification $\to$ v3.0
  \end{enumerate}
  \end{column}
  \begin{column}{0.5\textwidth}
  \textbf{Part III --- Paper A (current focus)}
  \begin{enumerate}\setcounter{enumi}{8}\small
    \item Story: memory as a routed interface
    \item Core results
    \item Real long-context transfer
    \item Full experiment status
  \end{enumerate}
  \vspace{0.6em}
  \textbf{Part IV --- Paper C (next line)}
  \begin{enumerate}\setcounter{enumi}{12}\small
    \item Why linear attention; the research line
    \item The landscape (delta-rule family)
    \item GDN forgetting: exploration \& modules
  \end{enumerate}
  \vspace{0.6em}
  \textbf{Part V --- Wrap}
  \begin{enumerate}\setcounter{enumi}{15}\small
    \item Development logic \& timeline
    \item Cross-cutting insights \& next steps
    \item Artifact index (all links)
  \end{enumerate}
  \end{column}
  \end{columns}
\end{frame}

% ============================ PART I ============================
\section{Framing}
\begin{frame}{1. Background: the long-context memory wall}
  \textbf{Problem.} Long inputs carry documents, tools, history, and unfinished agent work, but processing
  every token requires more compute and memory as context grows.
  \begin{itemize}
    \item A larger window does not guarantee that the model uses the needed evidence; position,
    distractors, and reasoning depth still change the answer.
    \item \textbf{Two practical responses:}
    \begin{enumerate}
      \item \emph{Select raw evidence} --- retrieval, prompt compression, or a bounded window.
      \item \emph{Build a compact state} --- learned soft memory or a recurrent linear-attention state.
    \end{enumerate}
    \item \textbf{System requirement:} preserve a raw path when the compact state is insufficient.
  \end{itemize}
  \vspace{0.4em}
  \begin{block}{Guiding question}
  On a \emph{frozen} base, what should be kept as compact memory, and when should an agent return to raw evidence?
  \end{block}
\end{frame}

\begin{frame}{2. One thesis, three papers (the mainline)}
  \begin{block}{Thesis}
  Long-context memory is an \best{interface and routing} problem: a compact state is useful only where
  the next decision survives compression, and raw evidence must remain available outside that region.
  \end{block}
  \begin{center}
  \resizebox{\textwidth}{!}{%
  \begin{tikzpicture}[
     box/.style={draw, rounded corners, thick, minimum height=1.9cm, text width=3.0cm, align=center, font=\scriptsize},
     lbl/.style={midway, above, font=\tiny, text=black!70},
     node distance=2.1cm]
    \node[box, fill=blue!8]  (a) {\textbf{Paper A}\\[2pt] recurrent soft memory\\ {\small +} raw fallback\\[2pt] \textcolor{blue!65!black}{\tiny current focus}};
    \node[box, fill=green!12, right=of a] (b) {\textbf{Paper B}\\[2pt] input-side selection\\ {\small +} IMP diagnosis\\[2pt] \textcolor{green!45!black}{\tiny training-free line}};
    \node[box, fill=orange!15, right=of b] (c) {\textbf{Paper C}\ (v3.0)\\[2pt] linear attention\\ {\small \&} GDN forgetting\\[2pt] \textcolor{orange!75!black}{\tiny next line}};
    \draw[-{Latex}, thick] (a) -- (b) node[lbl]{latent state $\to$ raw selection};
    \draw[-{Latex}, thick] (b) -- (c) node[lbl]{selection on linear};
  \end{tikzpicture}}
  \end{center}
  \vspace{0.2em}
  \footnotesize Paper A asks when a learned short state is usable; Paper B asks which raw evidence should
  be retained without training; Paper C studies how the same selection problem changes for recurrent
  linear-attention memory.
\end{frame}

% ============================ PART II: Paper B ============================
\section{Paper B --- diagnosis + IMP}
\begin{frame}{3. Paper B --- the diagnosis (model-invariant)}
  Systematic full-test map across 9 model families (1.7B--14B, Qwen2.5/3/3.5, GLM-4, Ministral, Llama; quad + linear):
  \begin{itemize}
    \item \best{No universal compressor} --- the winner flips by task (retrieval / extractive / MC / reasoning).
    \item \best{``More context can hurt''} --- on literary-MC \emph{full} $<$ blind guessing (QuALITY full 7.2 $<$ 17.9); compression \emph{denoises}.
    \item \best{Retrieval $\gg$ full} on ultra-long ($\infty$Bench) --- selecting evidence beats reading everything.
    \item \best{Model-invariant}: the crossover \& denoising hold on \emph{every} family, incl. cache-free linear GDN.
  \end{itemize}
  \begin{block}{}
  \footnotesize This diagnosis --- not a single number --- is Paper B's durable contribution.
  \ghB{matrix-facts.md}{fact base (F27/F31/F43)} \; \ghB{main-table-fulltest.md}{full-test main table}
  \end{block}
\end{frame}

\begin{frame}{4. Paper B --- the method: IMP (Importance-routing Prefilter)}
  \textbf{Training-free, input-side, architecture-agnostic} prefilter on a \emph{frozen} base:
  \begin{enumerate}\small
    \item Score context tokens/chunks by cheap $O(L)$ signals (lexical BM25 + query-relevance + surprisal).
    \item Keep the top spans/chunks \emph{verbatim} to a budget; drop the rest. Feed to the frozen base.
    \item \best{Route} the granularity per input (\texttt{auto}): long/extractive $\to$ chunk-BM25 over the \emph{full doc}; literary-MC-that-fits $\to$ token-importance spans.
  \end{enumerate}
  \begin{itemize}\small
    \item No weights trained; runs on dense \emph{and} linear bases; no KV cache needed.
    \item Covers regimes that RAG (fails reasoning) and KV-eviction (can't run on linear) each only half-cover.
  \end{itemize}
  \begin{block}{}
  \footnotesize \ghB{method-vNext-2026-07-13.md}{method spec + code path} \; \ghB{imp-method-and-implementation.md}{implementation}
  \end{block}
\end{frame}

\begin{frame}{5. Paper B --- full results (Qwen3-8B, all methods $\times$ benches)}
  \centering\scriptsize
  \begin{tabular}{lccccccc}
  \toprule
  bench & full$\cdot$t & rag & kvzip$\ddagger$ & knorm$\ddagger$ & tome & \best{auto (ours)} & best \\
  \midrule
  $\infty$Bench (>100k) & 52.8 & 64.2 & 35.8 & 25.8 & 51.5 & \best{70.3} & \best{ours} \\
  RULER-NIAH & 99 & 94 & \best{99} & 85 & 0 & 98 & kvzip/ours* \\
  squad\_v2 & 65.9 & 67.2 & 64.9 & 16.5 & 38.9 & 66.9 & rag$\approx$ours \\
  trivia\_qa & 70.9 & 72.7 & 71.2 & 40.7 & 62.3 & \best{72.8} & \best{ours} \\
  lb\_multifieldqa & 38.6 & 38.8 & 37.1 & 30.7 & -- & \best{39.5} & \best{ours} \\
  lb\_narrativeqa & 14.0 & 13.7 & 11.2 & 9.3 & 13.5 & \best{14.3} & \best{ours} \\
  hotpot\_qa & 57.2 & 55.8 & 56.2 & 23.1 & 43.9 & 48.8 & full \\
  QuALITY-hard & 9.2 & 10.5 & \best{29.9} & 29.6 & 11.7 & 12.9 & \best{kvzip$\ddagger$} \\
  MuSR & 58.9 & 56.7 & 55.6 & 53.3 & 48.9 & 48.9 & full \\
  \bottomrule
  \end{tabular}
  \vspace{0.3em}
  \footnotesize $\cdot$t = truncated to 16k; $\ddagger$ = needs KV cache (\emph{cannot} run on linear). ours = \texttt{auto}.
  \ghB{COMPLETE-RESULTS-2026-07-13.md}{complete table (16 benches, 9 families)}
\end{frame}

\begin{frame}{6. Paper B --- honest standing vs \emph{all} baselines (not just RAG)}
  \begin{columns}[T]
  \begin{column}{0.52\textwidth}
  \textbf{Where ours wins / ties}
  \begin{itemize}\small
    \item \best{Wins}: $\infty$Bench, trivia, multifieldqa, narrativeqa.
    \item \best{Ties best}: squad, lb\_hotpotqa, ms\_marco, LongBench-v2; best \emph{arch-agnostic} on RULER.
    \item \best{Only method that runs on linear} among the strong ones.
  \end{itemize}
  \end{column}
  \begin{column}{0.48\textwidth}
  \textbf{Where we honestly lose}
  \begin{itemize}\small
    \item \textcolor{red}{Literary-MC} (QuALITY) $\to$ \emph{KV-eviction} (kvzip/knorm $\approx$30) --- but they need a KV cache.
    \item \textcolor{red}{Dense reasoning} (MuSR, musique) $\to$ \emph{full} (needs everything).
  \end{itemize}
  \end{column}
  \end{columns}
  \vspace{0.4em}
  \begin{block}{Framing for the paper}
  \footnotesize Contribution = the \best{diagnosis} + a \best{regime-spanning training-free router}, \emph{not} ``beats everything''.
  We state plainly where KV-eviction / full win. \ghB{best-of-all-baselines-2026-07-13.md}{best-of-all comparison}
  \end{block}
\end{frame}

\begin{frame}{7. Paper B --- rigor: a bug we found \& fixed}
  \begin{itemize}
    \item \textbf{Symptom:} IMP looked $-19$ vs RAG on $\infty$Bench.
    \item \textbf{Root cause (case-level audit):} IMP \emph{truncated} the context to 16k \emph{before} selecting; RAG retrieved over the \emph{whole} 131k-token doc. \best{87\% of RAG's kept content came from beyond 16k --- IMP was structurally blind to it.}
    \item \textbf{Fix:} chunk-family retrieves over the full doc (no forward needed). $\infty$Bench $51 \to \best{76}$ --- the gap \emph{reversed} to a win.
    \item \textbf{Discipline installed:} every result labels \texttt{full$\cdot$trunc16k}; per-bench truncated-vs-true-full audited; only $\infty$Bench materially affected.
  \end{itemize}
  \begin{block}{}
  \footnotesize \ghB{correctness-audit-2026-07-13.md}{correctness audit} \; \ghB{imp-vs-rag-case-diff-2026-07-13.md}{case-level diff} \; \ghB{code-review-2026-07-13.md}{code review}
  \end{block}
\end{frame}

\begin{frame}{8. Paper B --- Occam simplification $\to$ v3.0}
  \textbf{What actually drives the wins?} An ablation of the (baroque) \texttt{auto} router shows:
  \begin{itemize}\small
    \item The \best{chunk branch = full-doc BM25 retrieval} carries nearly all the wins.
    \item Importance signals (surprisal, query-dot) add $\approx 0$; the span branch only matters in literary-MC (where it still loses to KV-eviction).
  \end{itemize}
  \begin{block}{v3.0 = \texttt{IMP-lite}: one mode, one signal, one knob}
  \footnotesize Full-document lexical chunk retrieval on a frozen base --- no span mode, no surprisal, no router, no training.
  Occam: drop every non-load-bearing part; keep the workhorse. \ghB{method-vNext-2026-07-13.md}{v3.0 design + validation plan}
  \end{block}
  \footnotesize \emph{Open question (decide empirically):} is IMP-lite meaningfully $>$ RAG, or is it ``RAG configured for compression + the diagnosis''?
\end{frame}

% ============================ PART III: Paper A ============================
\section{Paper A}
\begin{frame}{9. Paper A --- memory is a routed interface}
  \begin{block}{Question}
    When can an agent answer from a short learned state, and when should it return to raw evidence?
  \end{block}
  \begin{center}
  \resizebox{0.98\textwidth}{!}{%
  \begin{tikzpicture}[
    node distance=0.55cm,
    box/.style={draw, rounded corners, minimum height=0.9cm, text width=2.15cm, align=center, font=\scriptsize},
    arr/.style={-{Latex}, thick}]
    \node[box, fill=black!4] (ctx) {context chunks\\+ query};
    \node[box, fill=blue!10, right=of ctx] (enc) {frozen LM\\recurrent encoder};
    \node[box, fill=green!12, right=of enc] (mem) {trainable projection\\soft memory};
    \node[box, fill=blue!10, right=of mem] (read) {same frozen LM\\read LoRA on};
    \node[box, fill=purple!10, right=of read] (gate) {confidence router};
    \node[box, fill=orange!15, right=of gate] (ans) {final answer};
    \node[box, fill=black!4, below=0.55cm of read] (raw) {bounded raw path\\read LoRA off};
    \draw[arr] (ctx)--(enc); \draw[arr] (enc)--(mem); \draw[arr] (mem)--(read);
    \draw[arr] (read)--(gate); \draw[arr] (gate)--(ans);
    \draw[arr] (ctx.south) |- (raw.west); \draw[arr] (raw.east) -| (gate.south);
  \end{tikzpicture}
  }
  \end{center}
  \begin{itemize}\small
    \item One shared backbone supplies encoding, memory reading, confidence, and raw fallback.
    \item The contribution is the \best{complete compact-versus-raw route}, not compression ratio alone.
    \item The formal fixed-threshold test currently certifies no nonzero coverage; routing results are held-out empirical analysis.
  \end{itemize}
  \begin{block}{}\footnotesize \ghA{PAPER-A-draft-v0.1.md}{paper story} \; \ghA{paper-a-all-benchmarks-skeleton-2026-07-21.md}{all-benchmark tables}\end{block}
\end{frame}

\begin{frame}{10. Paper A --- core results (three seeds)}
  \centering\scriptsize
  \begin{tabular}{llrrr}
  \toprule
  base & method & QuALITY & BFCL & HotpotQA \\
  \midrule
  \multirow{6}{*}{Qwen3-8B}
    & \refm{Raw} & 7.2 & 92.4 & 53.7 \\
    & \refm{SFT} & 81.7$\pm$1.7 & 95.4$\pm$0.3 & 68.8$\pm$0.6 \\
    & Window & 15.7 & 55.7 & 26.2 \\
    & LLMLingua-2 & 14.3 & 70.3 & 22.1 \\
    & GCM & 54.4$\pm$0.2 & 72.3$\pm$0.5 & 28.9$\pm$0.2 \\
    & \best{Route} & \best{54.6} & \best{88.5} & \best{50.9} \\
  \midrule
  \multirow{6}{*}{Qwen3.5-9B}
    & \refm{Raw} & 7.1 & 84.5 & 53.9 \\
    & \refm{SFT} & 85.0$\pm$0.4 & 94.9$\pm$1.0 & 71.7$\pm$0.6 \\
    & Window & 16.7 & 52.8 & 24.8 \\
    & LLMLingua-2 & 20.3 & 60.8 & 28.9 \\
    & \best{GCM} & \best{51.5$\pm$1.7} & 72.0$\pm$0.8 & 30.5$\pm$0.3 \\
    & \best{Route} & 51.4 & \best{80.5} & \best{51.7} \\
  \bottomrule
  \end{tabular}
  \vspace{0.35em}
  \begin{itemize}\footnotesize
    \item GCM is the strongest completed compact path across the displayed core tasks; SFT is a full-cost reference.
    \item Routing recovers much of the raw gap on BFCL and HotpotQA; QuALITY already favors compact memory.
    \item SQuAD-v2 is retained only as a short-passage exact-text diagnostic, not a headline benchmark.
  \end{itemize}
  \begin{block}{}\footnotesize \ghA{current-data-2026-07-19/main-table.md}{audited core table} \; \ghA{current-data-2026-07-19/gate-summary.md}{routing metrics}\end{block}
\end{frame}

\begin{frame}{11. Paper A --- real long-context transfer (seed 42)}
  \centering\scriptsize
  \begin{tabular}{llrrr}
  \toprule
  base & method & LongBench-v2 & $\infty$Bench-choice & BABILong \\
  \midrule
  \multirow{4}{*}{Qwen3-8B}
    & \refm{No context} & 33.4 & 41.9 & 2.0 \\
    & \refm{Raw} & 31.2 & 52.8 & 18.1 \\
    & GCM & 20.1 & 21.4 & 0.9 \\
    & \best{Route} & \best{31.4} & \best{52.8} & \best{18.1} \\
  \midrule
  \multirow{4}{*}{Qwen3.5-9B}
    & \refm{No context} & 27.0 & 42.4 & 3.1 \\
    & \refm{Raw} & 33.0 & 52.4 & 10.5 \\
    & GCM & 22.5 & 31.9 & 2.2 \\
    & \best{Route} & \best{33.4} & \best{52.4} & \best{11.9} \\
  \bottomrule
  \end{tabular}
  \vspace{0.35em}
  \begin{block}{New result: the boundary is task-structured}
    \footnotesize On multi-hop LongBench transfer, GCM can beat truncated raw
    (Qwen3 HotpotQA: $19.1\to30.5$; Qwen3.5: $3.9\to26.6$), and Route improves further.
    On single-document / extractive targets, raw wins and routing falls back. Across the harvested
    two-base targets, \best{Route $\ge$ Raw in every cell}.
  \end{block}
  \footnotesize Single seed; error bars pending. RULER and $\infty$Bench K32 repairs remain.
  \begin{block}{}\footnotesize \ghA{paper-a-all-benchmarks-skeleton-2026-07-21.md}{full long-context breakdown}\end{block}
\end{frame}

\begin{frame}{12. Paper A --- full execution status (Wed 2026-07-22 06:33 PT)}
  \begin{columns}[T]
  \begin{column}{0.52\textwidth}
    \centering\scriptsize
    \begin{tabular}{lrrrr}
    \toprule
    stage & done & run & fail & pending \\
    \midrule
    core main & 88 & 0 & 0 & 0 \\
    replicate + SFT audit & 9 & 0 & 0 & 0 \\
    transfer adapters & 46 & 0 & 2 & 0 \\
    real long-context & 50 & 3 & 2 & 63 \\
    8-model generality & 11 & 1 & 0 & 36 \\
    budget / RULER & 0 & 0 & 0 & 23 \\
    ablation & 0 & 0 & 0 & 36 \\
    official baselines & 0 & 0 & 0 & 52 \\
    \bottomrule
    \end{tabular}
    \vspace{0.45em}
    \begin{block}{Live workers}
      \scriptsize Three Qwen3.5-4B long-context jobs + one Qwen3.5-9B generality job.
      One GPU on the generality node is unavailable.
    \end{block}
  \end{column}
  \begin{column}{0.48\textwidth}
    \textbf{Completed evidence}
    \begin{itemize}\scriptsize
      \item Core grid, output audit, gate analysis, replicate, SFT reaudit.
      \item Two-base long-context harvest: 44 cells.
    \end{itemize}
    \textbf{Repairs / blockers}
    \begin{itemize}\scriptsize
      \item K32 $\infty$Bench adapters failed on both main bases.
      \item GLM-4 and Qwen3.5-9B K32 source-adapter repairs remain.
      \item LCLM / Semi-Dynamic cloned; official execution not started.
    \end{itemize}
    \textbf{Execution order}
    \begin{enumerate}\scriptsize
      \item continue long-context across the remaining bases;
      \item complete generality;
      \item budget/RULER $\to$ ablation/cost;
      \item official baselines.
    \end{enumerate}
  \end{column}
  \end{columns}
  \begin{block}{}\footnotesize \ghA{paper-a-full-benchmark-table-2026-07-21.md}{benchmark + baseline inventory} \; \ghA{experiment-receipt-2026-07-16.md}{experiment receipt}\end{block}
\end{frame}

% ============================ PART IV: Paper C ============================
\section{Paper C --- linear attention}
\begin{frame}{13. Paper C (v3.0) --- why linear attention; the research line}
  \textbf{Focus (next line):} linear attention \& its variants (Qwen3.5/3.6 GDN family; \emph{no} Qwen3 control).
  \begin{itemize}\small
    \item \textbf{The tension:} a \emph{fixed-size state} is efficient but has a structural \best{recall / forgetting bottleneck} (recall--throughput tradeoff; state-tracking limits).
    \item \textbf{Our hook (already shown):} on Qwen3.5-GDN the long-context diagnosis reproduces, but \best{KV-eviction can't even run} (no KV) --- only input-side selection (IMP-lite/RAG) works.
    \item \textbf{The opening:} on linear bases the problem isn't ``shrink the KV'' (there is none) --- it's \best{feed the right small context / augment the state's recall}.
  \end{itemize}
  \begin{block}{Candidate spine}
  \footnotesize C-1 diagnose GDN forgetting (no training) + C-2 training-free input-side memory (IMP-lite) that KV-methods structurally can't provide. \ghC{PAPER-C-research-line.md}{research line}
  \end{block}
\end{frame}

\begin{frame}{14. Paper C --- the landscape (delta-rule family)}
  \centering\scriptsize
  \begin{tabular}{lcccc}
  \toprule
  model & decay & erase & write & delta \\
  \midrule
  Mamba-2 & scalar & --- & scalar & no \\
  \textbf{Gated DeltaNet} (our base) & scalar & tied scalar & tied scalar & yes \\
  KDA / Kimi-Linear & channel-wise & tied scalar & tied scalar & yes \\
  \textbf{GDN-2} (NVIDIA, 2026) & channel-wise & \best{channel-wise} & \best{channel-wise} & yes \\
  \bottomrule
  \end{tabular}
  \vspace{0.5em}
  \begin{itemize}\footnotesize
    \item Frontier = \best{how to \emph{edit} the compressed memory} (erase vs write) without scrambling associations.
    \item Field findings: the \emph{delta rule} dominates recall; \best{GDN-2's erase gate carries most of its gain} (selective forgetting of stale keys is the lever); channel-wise helps long-range but can hurt precise recall.
  \end{itemize}
  \begin{block}{}\footnotesize \ghC{references-linear-attention.md}{reference collection (11 categories, arXiv)}\end{block}
\end{frame}

\begin{frame}{15. Paper C --- GDN forgetting: exploration \& modules}
  \textbf{Training-free anti-forgetting modules tested on Qwen3.5-GDN (real benches):}
  \centering\scriptsize
  \begin{tabular}{lccccc}
  \toprule
  bench & full & imp-lite & rel-last & replay & rag \\
  \midrule
  $\infty$Bench & 55 & 76 & 75 & \best{78} & 65 \\
  lb\_multifieldqa & 38.6 & 42.1 & 40.7 & \best{44.0} & 41.1 \\
  babilong-qa1 & 65 & \best{68} & 62 & 61 & 54 \\
  \bottomrule
  \end{tabular}
  \vspace{0.4em}
  \begin{itemize}\footnotesize
    \item \best{Input-side compression rescues GDN on ultra-long} ($\infty$Bench 55$\to$76) --- don't overflow the fixed state.
    \item Recency-reorder modules (rel-last / replay) = \emph{marginal, task-dependent} $\Rightarrow$ Occam: imp-lite default.
    \item \textcolor{red}{Honesty:} our synthetic single-needle probe was \emph{inconclusive} (task too easy / a probe bug we caught); the trustworthy evidence is the \emph{real-bench} grid above. 128k probe OOMs (GDN kernel).
  \end{itemize}
  \begin{block}{}\footnotesize \ghC{gdn-forgetting-exploration.md}{ablation design} \; \ghC{gdn-module-results-2026-07-14.md}{module results (+ probe caveats)}\end{block}
\end{frame}

% ============================ PART V: Wrap ============================
\section{Wrap}
\begin{frame}{16. Development logic \& timeline}
  \begin{enumerate}\small
    \item \textbf{Paper A} --- recurrent soft memory + raw fallback: core grid complete; two-base long-context transfer harvested; remaining scale, budget, ablation, and official baselines running.
    \item \textbf{Paper B} --- complementary input-side diagnosis + IMP selection on a frozen base; full-test audit and Occam simplification completed.
    \item \textbf{Shared result} --- no representation wins everywhere; task structure determines whether compact memory, selected raw spans, or bounded raw context should answer.
    \item \textbf{Paper C} --- carry the selection question to recurrent linear-attention memory, where state forgetting replaces raw-window growth as the main bottleneck.
    \item \textbf{Current priority} --- finish Paper A's remaining grids and lock its complete result tables before expanding Paper C.
  \end{enumerate}
\end{frame}

\begin{frame}{17. Cross-cutting insights \& next steps}
  \textbf{Insights}
  \begin{itemize}\small
    \item Compact memory has a \best{task-structured operating region}: strong on some decision and multi-hop interfaces, weak on exact planted evidence.
    \item \best{No universal context path}; the agent should choose among learned memory, selected raw evidence, and bounded raw context.
    \item Reliability must be evaluated jointly with coverage and total cost, not only average compressed accuracy.
  \end{itemize}
  \textbf{Next steps}
  \begin{itemize}\small
    \item \textbf{Paper A:} complete transfer / generality; run RULER, budget, ablation, cost, and official-native baselines; then lock tables and figures.
    \item \textbf{Paper B:} run the IMP-lite Occam ablation; finalize v3.0; decide the RAG-delta claim.
    \item \textbf{Paper C:} secure the in-band \best{Qwen3.6-27B} (hybrid) checkpoint; RULER access (HF); diagnose GDN forgetting on real long-context; test input-side vs erase-gate-retrofit fixes.
  \end{itemize}
\end{frame}

\begin{frame}{18. Artifact index (all links to GitHub)}
  \scriptsize
  \textbf{Paper B} \quad
  \ghB{REPORT-2026-07-13.md}{consolidated report} $\cdot$
  \ghB{COMPLETE-RESULTS-2026-07-13.md}{complete results} $\cdot$
  \ghB{main-table-fulltest.md}{main table} $\cdot$
  \ghB{matrix-facts.md}{fact base} $\cdot$
  \ghB{best-of-all-baselines-2026-07-13.md}{best-of-all} \\
  \ghB{correctness-audit-2026-07-13.md}{correctness audit} $\cdot$
  \ghB{code-review-2026-07-13.md}{code review} $\cdot$
  \ghB{method-vNext-2026-07-13.md}{v3.0 method} $\cdot$
  \ghB{FIVE-DAY-SUMMARY-2026-07-08_13.md}{5-day summary}
  \vspace{0.7em}

  \textbf{Paper A} \quad
  \ghA{PAPER-A-draft-v0.1.md}{paper draft} $\cdot$
  \ghA{paper-a-all-benchmarks-skeleton-2026-07-21.md}{all tables} $\cdot$
  \ghA{paper-a-full-benchmark-table-2026-07-21.md}{benchmark inventory} $\cdot$
  \ghA{paper-baselines-v2-2026-07-19.md}{baseline plan} $\cdot$
  \ghA{experiment-receipt-2026-07-16.md}{experiment receipt}
  \vspace{0.7em}

  \textbf{Paper C} \quad
  \ghC{README.md}{overview} $\cdot$
  \ghC{PAPER-C-research-line.md}{research line} $\cdot$
  \ghC{references-linear-attention.md}{references} $\cdot$
  \ghC{gdn-forgetting-exploration.md}{GDN forgetting design} $\cdot$
  \ghC{gdn-module-results-2026-07-14.md}{module results}
  \vspace{0.9em}

  \centering \normalsize Repo: \href{https://github.com/BDeMo/research-notes}{\color{blue}\underline{github.com/BDeMo/research-notes}}
\end{frame}

\end{document}
